### 2019-11-26 Mathematical Statistics/ms_ch_03.rmd
Mathematical Statistics, ch 3
Mathematical Statistics, ch 3

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
library(gtools, quietly = TRUE, warn.conflicts = FALSE)

@@

>>>>

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
library(gtools, quietly = TRUE, warn.conflicts = FALSE)

@@

### 3.12

In the Iowa recidivism case study in Section 1.4, offenders had originally been convicted of either a felony or misdemenor.

```{r, echo = T}
Recidivism <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
                               header = T))
```

a.) Use R to create a table displaying the proportion of felons who recidivated and the propportion of those convictted of a misdemeanor who recidivated.

Recidivism[, Recidivated := is.na(Days)]

Recidivated <- Recidivism[, .(RecidPct = sum(Recidivated) / .N), by = Offense]

pretty_kable(Recidivated, "Recidivated")

b.) Determine whether or not the difference in recidivism proportions computed in (a) is statistically significant.

observed <- as.numeric( Recidivated[Offense == "Felony"]$RecidPct - Recidivated[Offense == "Misdemeanor"]$RecidPct )

for( i in 1:N )
{
   index <- sample(nrow(Recidivism), nrow(Recidivism[Offense == "Felony"]), replace = F)
   results[i] <- as.numeric( Recidivism[index, .(Pct = sum(Recidivated) / .N)] - Recidivism[-index, .(Pct = sum(Recidivated) / .N)] )
}

p <- ( sum(results >= observed) + 1 ) / ( N + 1 )

It appears there is statistical significance in the recidivism rates between felony and misdemeanor convictions.

P-value: `r p`

### 3.13

In the Iowa recidivism case study in Section 1.4, for those offenders who recidivated, we have data on the number of days until they reoffended.

For those offenders who did recidivate, determine if the difference in the mean number of days (__Days__) until recidivism between those under 25 years of age and those 25 years of age and older is statistically significant.

Recid.Age25 <- Recidivism[, .(RecidDays = mean(Days, na.rm = T)), by = Age25]

pretty_kable(Recid.Age25, "Days Until Recidivism by Age ( < 25 < )")

observed <- as.numeric( Recid.Age25[Age25 == "Under 25"]$RecidDays - Recid.Age25[Age25 == "Over 25"]$RecidDays)

for( i in 1:N )
{
   index <- sample(nrow(Recidivism), nrow(Recidivism[Age25 == "Under 25"]), replace = F)
   results[i] <- mean(Recidivism[index]$Days, na.rm = T) - mean(Recidivism[-index]$Days, na.rm = T)
}

p <- min(1, ( sum(results >= observed) + 1 ) / ( N + 1 ) )

There does not appear to be a statistical difference between the days until recidivism by those over/under 25 years of age.

P-value: `r p`

### 3.14

Does chocolate ice cream have more calories than vanilla ice cream? The data set __IceCream__ contains calorie information for a sample of brands of chocolate and vanilla ice cream.

```{r, echo = T}
IceCream <- data.table(read.csv(paste0(data.dir, "IceCream.csv"),
                               header = T))
```

a.) Inspect the data set, then explain why this is an example of matched pairs data.

```{r, echo = T}
pretty_kable(IceCream, "Ice Cream Data")
```

This data set is matched pairs becase there are multiple brands associated with each flavor. They are not independent samples.

b.) Compute summary statistics of the number of calories for the two flavors.

pretty_kable(IceCream[, .(Vanilla = mean(VanillaCalories), Chocolate = mean(ChocolateCalories))], 
             "Ice cream Calories by Flavor")

c.) Conduct a permutation tests to determine whether or not chocolate ice cream has, on average, more calories than vanilla ice cream.

diff <- IceCream$VanillaCalories - IceCream$ChocolateCalories
observed <- mean(diff)

for( i in 1:N )
{
   Sign <- sample(c(-1, 1), nrow(IceCream), replace = T)
   diff2 <- Sign * diff
   results[i] <- mean(diff2)
}

p <- ( sum( results <= observed ) + 1 ) / ( N + 1 )

The data supports the claim that chocolate ice cream has more calories than vanilla.

P-value: `r p`

### 3.15

Is there a difference in the price of groceries sold by the two retailers Target and Walmart?

The data set Groceries contains a sample of grocery items and their prices advertised on their respective web sites on one specific day.

```{r, echo = T}
Groceries <- data.table(read.csv(paste0(data.dir, "Groceries.csv"),
                               header = T))

a.) Inspect the data set, then explain why this is an example of matched pairs data.

```{r, echo = T}
pretty_kable(Groceries, "Target vs Walmart")
```

b.) Compute summary statistics of the prices for each store.

diff <- Groceries$Target - Groceries$Walmart

observed <- mean(diff)

Observed difference: __`r observed`__

c.) Conduct a permutation test to determine whether or not there is a difference in the mean prices.

```{r, echo = T}
N <- 10e2 - 1
results <- numeric(N)

for( i in 1:N )
{
   Sign <- sample(c(-1, 1), nrow(Groceries), replace = T)
   diff2 <- Sign * diff
   results[i] <- mean(diff2)
}

p <- min(1, 2 * ( sum(results >= observed ) + 1 ) / ( N + 1) )

There does not appear to be a statistical difference in the prices at the two stores.

P-value: `r p`

d.) Create a histogram of the difference in prices.

```{r, echo = T, fig.height=5, fig.width=8}

ggplot(data.table( Product = Groceries$Product, Diff = diff)) +
   geom_bar(aes(Product, Diff, fill = Product), stat = "identity") +
   coord_flip() +
   labs(title = "Grocery Price Difference (Target vs Walmart)") +
   theme(legend.position = "none")

What is unusual about Quaker Oats Life cerial?

__Price difference is a multiple SD outlier.__

e.) Redo the hypothesis test without this observation.

diff <- Groceries[Product != "Quaker Oats Life Cereal  Original ", .(Diff = Target - Walmart ) ]$Diff

observed <- mean(diff)

for( i in 1:N )
{
   Sign <- sample(c(-1, 1), length(diff), replace = T)
   diff2 <- Sign * diff
   results[i] <- mean(diff2)
}

p <- min(1, 2 * ( sum(results >= observed ) + 1 ) / ( N + 1) )

Do you reach the same conclusion?

__No, there is a statistical difference in the prices of products at the two retailers with the outlier removed.__

P-value: `r p`

### 3.16

In the sampling version of permutation testing, the one-sided P-value is

$\hat{P} = \frac{(X + 1)}{(N + 1)}$, where X is the number of permutation test statistics that are as large or larger than the observed test statistic.

Suppose the true P-value (for the exaustive test, conditional on the observed data) is __p__.

a.) What is the variance of $\hat{P}$?

$\mathbb{V} = p*(1 - p) / (N + 1)$

b.) What is the variance of $\hat{P}_2$ for the two-sided test (assuming that __p__ is not close to 0.5, where __p__ is the smaller true one-side P-value)?
